FIXME

### 2013-02-10 stackExchange/writeups/iccm-2013/writeup.tex
foo
Former-commit-id: 6026de11fa4b5c677a1252080980ce5fe850852a [formerly 890ccfc28319a6a92fb9af040c27e71992fdc0f9]
Former-commit-id: 15bb22acebf8dcb085d2ea3dc4b786dd9bead4a6

It turns out that there are many connections between ACT-R's declarative memory retrieval equations and other non-ACT-R information retrieval theories,
and many of those theories use a combination of weighting schemes for contextual relevance and global importance of a particular chunk in context.
For example, \citeA{Dumais1991} shows that a measure of how likely term ${_j}$ is in document ${_i}$ is often the product of a local weighting $L_{ji}$ and global weighting $G_{j}$ measure.
This idea maps rather nicely to ACT-R's contextual weighting.
That is, terms are context words ${_j}$, documents are tags ${_i}$, local weighting is strength of association $S_{ji}$, and global weighting is attentional weighting $W_{j}$.

\citeA{Dumais1991} states that predictions tend to improve when a global weighting scheme is used, so that words that occur frequently or with many documents (tags) are not weighted as high.
He reviews four different commonly-used weighting schemes; one of which is a scaled entropy measure.

@@

\subsection{PMI}

ACT-R's declarative memory retrieval equations appear in other information retrieval literature that use word co-occurrence counts, although it's usually called the Pointwise Mutual Information Index.
For example, \citeA{Fu2007} and \citeA{Farahat2004} have shown that ACT-R's retrieval equations are mathematically equivalent to the Pointwise Mutual Information (PMI) index when the sample size of the corpus is large.
\citeA{Budiu2007} and \citeA{Farahat2004} have shown that the PMI index for word co-occurrences is a strong predictor of word similarity and information scent.
And on large corpora \citeA{Budiu2007} has shown that PMI is the best predictor of word pair similarity (compared to LSA and generalized LSA).
It is interesting to note that ACT-R's declarative memory retrieval theory not only fits human performance data well, but is also a useful and competitive best predictor of word pair similarity in the general information retrieval literature that is less bounded by cognitive constraints.

@@

>>>>

It turns out that there are many connections between ACT-R's declarative memory retrieval equations and other non-ACT-R information retrieval theories,
and many of those theories use a combination of weighting schemes for contextual relevance and global importance of a particular chunk in context.
For example, \citeA{Dumais1991} shows that a measure of how likely term ${_j}$ is in document ${_i}$ is often the product of a local weighting $L_{ji}$ and global weighting $G_{j}$ measure.
This idea maps rather nicely to ACT-R's contextual weighting.
That is, terms are context words ${_j}$, documents are tags ${_i}$, local weighting is strength of association $S_{ji}$, and global weighting is attentional weighting $W_{j}$.
\citeA{Dumais1991} states that predictions tend to improve when a global weighting scheme is used, so that words that occur frequently or with many documents (tags) are not weighted as high.
He reviews four different commonly-used weighting schemes; one of which is a scaled entropy measure.

@@

@@

\subsection{Entropy and ACT-R's Attentional Weighting}

Model performance improved by using a scaled entropy measure to weight the importance of each contextual element in the body and title of a post.
However, it's uncommon for an ACT-R model that uses the declarative memory retrieval mechanisms to use anything other than equal attentional weighting for each chunk in context.
But it is much more common outside of the ACT-R modeling community to use both a local (strength of association) and global (attentional weight) term to predict document retrieval \citeA{Dumais1991}.
In fact, it's been shown that the Pointwise Mutual Information Index (PMI) (a commonly-used measure to predict document retrieval) is mathematically equivalent to ACT-R's retrieval equations when the sample size of the corpus is large (\citeNP{Budiu2007}, \citeNP{Farahat2004}).
Both global and local weightings are commonly used when the PMI measurement is used.
This suggests that using a global weighting term for chunks in context (e.g., the scaled entropy measure used ehre) alongside the more commonly-used local weighting (spreading activation) might be useful for other ACT-R models that are taking advantage of large stores of declarative memory datesets to build up the model's base knowledge.
